---
title: "SImulation Exercise"
author: "Mitch Harden"
output: html_document
---

This project is about sampling distribution of means for an exponential distribution of 1/lambda and the standard deviation is also also 1/lambda. lambda was set to 0.2 for all of the simulations. By simulating 1000 such distributions, it is shown that the  center of the distribution of means is `r mean.theo=1/lambda; mean.theo` with standard deviation `r sd.theo=1/lambda; sd.theo`.

#### Simulating the exponential RVs
```{r}
library(ggplot2)
set.seed(12345) ## Same as the combination on my luggage.
n <- 40
means <- sapply(1:nosim, function(dummy) {
     mean(rexp(n, lambda))
})
```

#### The mean and standard error of the sampling distribution of means
The mean and standard deviation of the distribution of averages was computed as `r meanDist = mean(means); meanDist` and `r sdDist = sd(means); sdDist`.  The standard deviation of the distribution of averages or standard error of the mean is `r se.theo = sd.theo/sqrt(n); se.theo`.  From these results, it can be seen that the computed values differ from theory by `r (meanDist-mean.theo)/mean.theo*100`% and `r (sdDist-se.theo)/se.theo*100`%.

#### Normality of the sampling distribution of means
Central limit theorem states that the sampling distribution of means approaches a normal distribution as $n \to \infty$. Graphically, it the asymptotic normality of the sampling distribution of means can be seen by looking at the following plot of the histogram with gaussian curve:
``` {r, echo=FALSE}
     library(ggplot2)
     dat <- data.frame(means = means)
     g <- ggplot(dat, aes(x = means)) 
     g <- g + geom_histogram(binwidth=.3, color="black", aes(y = ..density..))
     g <- g + stat_function(fun = dnorm, args=list(mean=mean.theo, sd=se.theo))
     g  
```

#### 95% confidence interval for $1/\lambda$
The 95% confidence interval for $1/\lambda$ is given by the equation $\bar{X} \pm 1.96 \frac{S}{\sqrt{n}}$.  The confidence interval for $1/\lambda$ is [`r confIntMean = meanDist + c(-1,1)*1.96*sdDist; confIntMean`]. $1/\lambda$ = `r mean.theo` is near the center as expected.

#### Conclusion
Sampling distribution of means asymptotically approaches normality with mean and standard deviation near expected values as predicted by the central limit theorem.