---
title: "Part2"
author: "Mitch Harden"
output: html_document
---

This report investigates the response of the growth of teeth at three dose levels of Vitamin C with each of two supplement types (orange juice or ascorbic acid).

## Exploratory Data Analysis and Data Summary
The ToothGrowth data set consists of the following variables and types:
``` {r, echo=FALSE}
     library(plyr)
     library(ggplot2)
     data(ToothGrowth)
     str(ToothGrowth)
```
This data set has `r sum(!complete.cases(ToothGrowth))` incomplete cases.  Additionally, each variable's summary statistics are
``` {r, echo=FALSE}
     summary(ToothGrowth)
```
Vitamin C dosage is a factor variable with the factors of `r unique(ToothGrowth$dose)`.

## Three Hypotheses to Test

1. No differnce between delivery by orange juice or absorbic acid overall.
2. No difference based on dosage level, regardless of delivery method.
3. No difference based on delivery method for each dosage.

### Hypothesis 1
The summary statistics are as follows:
```{r, echo=FALSE}
     statsQ1 <- ddply(ToothGrowth, .(supp), summarize, mean.len=mean(len),
                      var.len=var(len), num.obs=length(len))
     df.Q1 <- with(statsQ1,num.obs[1]+num.obs[2]-2)
     var.pool.Q1 <- with(statsQ1,
                    ((num.obs[1]-1)*var.len[1]+(num.obs[2]-1)*var.len[2])/df.Q1)
     t.Q1 <- with(statsQ1,
                  (mean.len[1]-mean.len[2])/
                       sqrt(var.pool.Q1*(1/num.obs[1]+1/num.obs[2])))
     alpha <- .05
     qt.Q1 <- qt(1-alpha/2, df = df.Q1)
     statsQ1
```
Using a two-sided t-test with $\alpha =$ `r alpha`, and since the computed t-value is greater than `r qt.Q1`( $t =$ `r t.Q1` ) we fail to reject the null hypothesis.

### Hypothesis 2
The summary statistics are as follows:
```{r, echo=FALSE}
     statsQ2 <- ddply(ToothGrowth, .(dose), summarize, mean.len=mean(len),
                      var.len=var(len), num.obs=length(len))
     df.Q2.12 <- with(statsQ2,num.obs[1]+num.obs[2]-2)
     df.Q2.23 <- with(statsQ2,num.obs[2]+num.obs[3]-2)
     df.Q2.13 <- with(statsQ2,num.obs[1]+num.obs[3]-2)     
     var.pool.Q2.12 <- with(statsQ2,
                    ((num.obs[1]-1)*var.len[1]+(num.obs[2]-1)*var.len[2])/df.Q2.12)
     var.pool.Q2.23 <- with(statsQ2,
                    ((num.obs[2]-1)*var.len[2]+(num.obs[3]-1)*var.len[3])/df.Q2.23)
     var.pool.Q2.13 <- with(statsQ2,
                    ((num.obs[1]-1)*var.len[1]+(num.obs[3]-1)*var.len[3])/df.Q2.13)
     t.Q2.12 <- abs(with(statsQ2,
                  (mean.len[1]-mean.len[2])/
                       sqrt(var.pool.Q2.12*(1/num.obs[1]+1/num.obs[2]))))
     t.Q2.23 <- abs(with(statsQ2,
                  (mean.len[2]-mean.len[3])/
                       sqrt(var.pool.Q2.23*(1/num.obs[2]+1/num.obs[3]))))
     t.Q2.13 <- abs(with(statsQ2,
                  (mean.len[1]-mean.len[3])/
                       sqrt(var.pool.Q2.13*(1/num.obs[1]+1/num.obs[3]))))
     qt.Q2.12 <- qt(1-alpha/2, df = df.Q2.12)
     qt.Q2.23 <- qt(1-alpha/2, df = df.Q2.23)
     qt.Q2.13 <- qt(1-alpha/2, df = df.Q2.13)    
     statsQ2
```
Using the same analysis as for Question 1, but for each of the three mean differences, we get the following two-sided t-scores:
```{r, echo=FALSE}
     results.Q2 <- data.frame(mean.diff = c("1-2", "2-3", "1-3"),
                              abs.t.score = c(t.Q2.12, t.Q2.23, t.Q2.13),
                              t.critical = c(qt.Q2.12, qt.Q2.23, qt.Q2.13))
     results.Q2
```
Since $|t|>t_{critical}$ for all three scenarios, we succssfully reject the null hypotheses.

### Hypothesis 3
The summary statistics are as follows:
``` {r, echo=FALSE}
     statsQ3 <- ddply(ToothGrowth, .(supp, dose), summarize, 
                    mean.len=mean(len), var.len=var(len), num.obs=length(len))
     df.Q3.14 <- with(statsQ3,num.obs[1]+num.obs[4]-2)
     df.Q3.25 <- with(statsQ3,num.obs[2]+num.obs[5]-2)
     df.Q3.36 <- with(statsQ3,num.obs[3]+num.obs[6]-2)     
     var.pool.Q3.14 <- with(statsQ3,
                    ((num.obs[1]-1)*var.len[1]+(num.obs[4]-1)*var.len[4])/df.Q3.14)
     var.pool.Q3.25 <- with(statsQ3,
                    ((num.obs[2]-1)*var.len[2]+(num.obs[5]-1)*var.len[5])/df.Q3.25)
     var.pool.Q3.36 <- with(statsQ3,
                    ((num.obs[3]-1)*var.len[3]+(num.obs[6]-1)*var.len[6])/df.Q3.36)
     t.Q3.14 <- abs(with(statsQ3,
                  (mean.len[1]-mean.len[4])/
                       sqrt(var.pool.Q3.14*(1/num.obs[1]+1/num.obs[4]))))
     t.Q3.25 <- abs(with(statsQ3,
                  (mean.len[2]-mean.len[5])/
                       sqrt(var.pool.Q3.25*(1/num.obs[2]+1/num.obs[5]))))
     t.Q3.36 <- abs(with(statsQ3,
                  (mean.len[3]-mean.len[6])/
                       sqrt(var.pool.Q3.36*(1/num.obs[3]+1/num.obs[6]))))
     qt.Q3.14 <- qt(1-alpha/2, df = df.Q3.14)
     qt.Q3.25 <- qt(1-alpha/2, df = df.Q3.25)
     qt.Q3.36 <- qt(1-alpha/2, df = df.Q3.36)    
     statsQ3
```
Repeating the same t-tests for the three mean differences, we get the following two-sided t-scores:
```{r, echo=FALSE}
     results.Q3 <- data.frame(dose = c("0.5", "1.0", "2.0"),
                              abs.t.score = c(t.Q3.14, t.Q3.25, t.Q3.36),
                              t.critical = c(qt.Q3.14, qt.Q3.25, qt.Q3.36))
     results.Q3
```
We fail to reject the null hypothesis only when the dosage is 2.0.

## Conclusion
First, there is no significant difference for tooth length by supplement type. Second, the dosage level does have a significant effect on tooth length. Third, when disregarding supplement method tooth length is significantly effected at the lower two dosages but not at the highest dose.